---
title: "Education, Child Bearing, and the Mother's Personal Income"
author: "Langwen Guan, Yuhang Ju, Zike Peng"
date: "October 19, 2020"
fontsize: 12pt
geometry: margin = 1in
header-includes:
   - \usepackage{setspace}
   - \doublespacing
includes:
  in_header: inheader.tex
output: pdf_document
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
library(data.table)
library(janitor)
library(tidyverse)
library(tinytex)
library(tufte)
library(stargazer)


raw_data <- read.csv('initial.csv')

fbirth <- as.numeric(raw_data$achb1c)

fbirth <- na.omit(fbirth)

income <- as.numeric(raw_data$ttlincg2)

income <- na.omit(income)

ed <- as.numeric(raw_data$ehg3_01b)

ed <- na.omit(ed)

ei <- ggplot(raw_data, aes(x = ed, y = income))

fi <- ggplot(raw_data, aes(x = fbirth, y = income))

fe <- ggplot(raw_data, aes(x = fbirth, y = ed))

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```
\newpage
## Abstract

This study documents an analysis of data from the Canadian General Social Survey of 2017. The dataset provides ample information regarding different factors which affect a family. Using these data, this study explores the relationship between the age at which Canadian women give birth to their first children, their levels of education, and their level of personal income. The study used a linear model to explain the relationships between these factors. The study is a contribution to the understanding of gender relations in the society. 

\newpage
## Introduction

Personal income is a very important factor in people’s lives. It enables one to be independent and to some extent free. Typically, this matter is more complex for women due to the issues of childbirth and education which impact personal income differently. These two factors also have a peculiar relationship with each other. Childbirth is one of the most important events in a woman’s life. In addition to bringing a young one to life, birth also changes various factors of the mother’s life. Some of these changes are temporary, and so they subside with time. For instance, weight gain which occurs after birth reverses shortly afterwards. However, other changes have a lasting impact. They include the academic performance and professional competitiveness of the new mothers. As such, they usually limit the set of opportunities available to them. Education usually has an opposite effect on a person’s professional and economic life. Education provides one with many opportunities to fulfill professional and financial needs. As a result, people who are more educated tend to generate a higher income and have a higher quality of life. This report documents a study which explains the relationship between the age of a mother when she gives birth to her first born, her education, and her income. It is significant to evaluate the opportunities presented to women and specifically mothers[@tinbergen1972;@waite1978]. The study seeks to answer these research questions:
1.	How does a woman's income vary with the age at which she got her first born?
2.	How does her education vary with the age at which she gave birth to her first born?
3.	How does her education impact her income?

## Data

The data used to study this relationship were obtained from the Canadian General Social Survey of 2017. The data were made from an interview or a questionnaire which were filled by individuals about themselves. This dataset contained a large amount of data about various aspects of family. In essence, the dataset had 20,602 rows of data divided into 80 categories. These categories included the respondents’ age, occupation, educational level, incomes, and more information about themselves, their partners, and their children. The respondents were residents in Canada aged between 15 and 80 years of age. This means that each one of them could provide reliable information about himself or herself. The dataset contained both qualitative and quantitative data since it was created to be as exhaustive as possible. It was then coded into .csv format in order to summarize it concisely. The information was coded appropriately in order to facilitate easy and effective analysis of the data. However, the data were limited to the Canadian demographic. They also lacked the aspect of change with time and were only useful for a cross section study of the variables of age at first birth, personal income, and education. The change in the variables with time is crucial to identify the progress the country is making towards supporting women as an effort towards achieving gender equality in all matters. Furthermore, the respondents were provided with choices in giving their responses. While this increased the efficiency of the research process, it made the analysis less effective. This is because the data appeared clustered, although they were not actually clustered in real life [@statistics2017]. 

## Model

In order to analyze this data, this study isolated these variables and formatted them in such a way to facilitate the creation of a linear model. After importing the dataset into R as a .csv file, the age at first birth dataset, named ``` raw_data$achb1c``` in the original file, was changed from a character vector into a numerical one to facilitate the creation of a model. The NAs were stripped from the data before defining the linear model. A linear model was chosen due to the direct relationships between the three variables. Furthermore, it better suited the hypothesized relationship between the variables involved in the study. The data were fitted to the linear model summarized in the table below[@hlavacmarek2018]. In order to understand the relationship between the independent variables, the study also included separate linear models which represented the relationship between education and personal income as well as between age at first birth and personal income. 

### Linear Model Representing the Relationship Between Personal Income, Age at First Birth, and Education
```{r modeljoint, results = "asis"}
modeljoint <- lm(income ~ fbirth + ed)
stargazer(modeljoint, title="model summary", type='latex', summary = TRUE)

```
The formula is:
\newline
income = -0.001466 fbirth + 0.005032 ed + 2.406185
\newline

### Linear Models Representing The Relationship Between Education and Age at First Birth
```{r model1, results = "asis"}
model1 <- lm(ed ~ fbirth)
stargazer(model1, title="model summary", type='latex', summary = TRUE)
```
As such, the formula for the model is:
\newline
ed = 0.004935 fbirth + 4.976032 
\newline
In making this model, the study did not split the income or education vector into males and females. The idea behind the unified vectors was that in Canada, gender equality is maintained in education. In personal income, the distribution is random between men and women. In essence, the study assumed that sex does not determine a person's income or education. After creating the model, the study cleaned the data and came up with a new dataset in .csv format. This dataset represented qualitative data. As such, it had column titles which were easier to relate with. This made the dataset essential in interpreting the data. While it is usually advisable to clean the data before creating the regression model, the cleaned dataset was in characters, making it difficult to analyze. 

## Results
The study discovered a direct relationship between both education and age at first birth; and personal income. As the graph of income against age at first birth depicts, these two variables have a direct negative relationship. This means that women who have children at a younger age generally make more money than those who have children later on in life. Conversely, education showed a direct positive relationship with personal income. This means that women who have achieved higher levels of education generally generate more income as compared to women who have not achieved education at such levels[@ronchetti1997]. 
 
```{r fig-margin, fig.margin=TRUE}
fi + geom_point()+ geom_smooth(method = "lm")
```
 
## Discussion

The study revealed a relationship between personal income, education, and age at first birth which has not been explored before. First, the gradient in the linear model of the relationship between education and personal income represented the unitary improvement in personal income with a unit increment in the educational level. The gradient was positive, implying that higher education leads to higher personal income. However, the gradient was very low. This means that higher education is not sufficiently rewarded. The study appreciates that this disparity may have originated from non-occupational income. For instance, investment, inheritance, and other supplementary income. Second, the study proved the hypothesis that women who have children at an early age get a lower personal income. However, the study could not prove a causal relationship between the data. As such, instead of having a child at an early age leading to a smaller personal income later on in life, the data may have shown that women who have high personal incomes prefer to have children at a young age. Third, the study revealed that there was no significant relationship between education and women's age at first birth. This means that generally, a woman's level of education does not influence or inform her decision to have a child at a young age or later on in life. Conversely, having a child at a young age or at a relatively advanced age does not influence a person's academic endeavors. 
This study was centered on women's educational progress, the age at which they gave birth to their first children, and their personal income, and the relationship between these variables. It established that the relationships between personal income and both the level of education and the age at first birth are direct and linear. Although this study was successful in creating a reliable body of knowledge, it was limited in how much data it could make use of. In future, it recommends that a longitudinal study be conducted to investigate the trend of these variables with time. The longitudinal study will be instrumental in determining whether Canadian women can maintain the balance they have between these three issues going into the future. Further, the study should look more closely to represent how these variables vary in different provinces in Canada. Indeed, many more studies should be conducted to establish whether women in different countries are able to navigate the relationship between their age at first birth, their educational level, and their personal income. This way, the issue of gender equality can be understood and tackled as the global problem that it is. This would pave way for the academic discourse towards creation of solutions which foster more holistic gender equality all around the world. The academic discourse would enable various authorities to implement policies which protect women and look out for them[@rao1988;@jokinen2012]. 

This code can be found on github in this public repository: <https://github.com/Juyuhang/Problem-Set-2.git>

\newpage
# References